% =============================================================================
\section{Concepts-Binding Coupling}
\label{sec:concepts}
% =============================================================================
Generic programming enables the implementation of generic algorithms,
which work on collections of different types, can be easily
maintained, extended, and customized, and are type safe and easier to
read. As mentioned in the introduction, \cgal{} rigorously adheres to
the generic programming paradigm. As a consequence, most components of
\cgal{} are either class or function templates. Many of the parameters
of these templates are described in terms of models of concepts. When
a class or a function template is instantiated, each one of its
template parameters is substituted with a model of one or more
concepts associated with the template parameter. Close to~750 concepts
can be identified in \cgal{} at the time this article is written. Most
hierarchy graphs of concepts are small. Few graphs, such as the graph
of concepts of the geometry traits of the \iiDArrangementsPackage{}
package, are quite large with intricate refinement relations. We use
clusters of closely related concepts to describe the refinement
relations among them; see, e.g., Figure~\ref{fig:atc}. We have introduced
tight coupling between concepts and (i) binding generations and (ii)
type annotation. We describe these relations in this section.

% -----------------------------------------------------------------------------
\subsection{Following Generic Concepts}
\label{sec:code:concepts}
% -----------------------------------------------------------------------------
For each concept we have introduced a template function that generates
bindings for all the type and function members required by the
concept. We call this function for every model of the concept. This
systematic approach guaranties that all the documented functions and
types, but nothing else, are exposed. Consider for example the cluster
of geometry traits concepts depicted in
Figure~\ref{fig:atc:central}. The template functions
\cppLstinline{export_AosBasicTraits\_2()},
\cppLstinline{export_AosXMonotoneTraits\_2()}, and
\cppLstinline{export_AosTraits\_2()} accept binding class
objects\footnote{A binding class object of type
  \cppLstinline{bp:class_<T>} is used to expose the C++ type
  \cppLstinline{T} to Python.} for particular geometry traits models
and populate it with all attributes that correspond to the
requirements of the concepts \cppLstinline{AosBasicTraits\_2},
\cppLstinline{AosXMonotoneTraits\_2}, and \cppLstinline{AosTraits\_2},
respectively. If a concept \cppLstinline{B} refines a concept
\cppLstinline{A}, then the function \cppLstinline{export_B()} calls
the function \cppLstinline{export_A()}. The bindings of different
traits models are generated in separate compilation units. As nodes in
the concept refinement graphs may have more than a single parent, we
ensure that a function that corresponds to a concept is not invoked
more than once in a given compilation unit; see Lines~3--4 in the
definition of the function \cppLstinline{export_AosBasicTraits\_2()}
in Listing~\ref{lst:cpp:exportBasic}.

\begin{lstfloat}
  \begin{lstlisting}[style=cppFramedStyle,basicstyle=\normalsize,%
                     label={lst:cpp:exportBasic},%
                     caption={Exposing attributes that correspond to the concept \cppLstinline{AosBasicTraits_2}.}]
template <typename GeometryTraits_2, typename ClassObject, typename Concepts>
void export_AosBasicTraits_2(ClassObject c, Concepts& concepts) {
  // Sentinel
  static bool exported = false;
  if (exported) return;

  // Expose attributes that correspond to the traits
  auto& classes = concepts.m_basic_traits_classes;
  classes.m_point_2 = bp::class_<Point_2>("Point_2")
    .def(bp::init<>())
    .def(bp::init<const Point_2&>());
  ~~$\cdots$~~
}
  \end{lstlisting}
\end{lstfloat}

Typically, a concept in our case requires the provision of nested
types that are themselves models of other concepts. For example, the
concept \cppLstinline{AosBasicTraits\_2} requires the provision of the
nested type \cppLstinline{Point_2} used to represent a two-dimensional
point. The type \cppLstinline{Point_2} must be
\cppLstinline{DefaultConstructible} and
\cppLstinline{CopyConstructible}. Let \cppLstinline{p} be an object of
type \cppLstinline{Point_2} required by the
\cppLstinline{AosBasicTraits\_2} concept. Most traits models also
support the calls \cppLstinline{p.x()} and \cppLstinline{p.y()}; they
return the $x$ and $y$ coordinates of the points $p$,
respectively. The number type used to represent the coordinates
depends on the traits instance; for example, the type
\cppLstinline{Kernel::FT} is used for the
\cppLstinline{Arr_segment_traits<Kernel>} traits instance. However,
the \cppLstinline{Point_2} type nested in the traits model
\cppLstinline{Arr_Bezier_curve_traits\_2} does not maintain an exact
representation of the coordinates. Thus, it instead supports the call
\cppLstinline{p.approximate()}, which returns a pair of floating point
numbers that only approximate the real coordinates. The function
\cppLstinline{export_AosBasicTraits\_2()} constructs a class object
for the \cppLstinline{Point\_2} type, but adds nothing but Python
attributes that expose the default and copy constructors of this type;
see Lines~10--11 in the definition of the function
\cppLstinline{export_AosBasicTraits\_2()} in the listing. It also
saves the class object of the \cppLstinline{Point_2} type (see
Line~8), so that later on the functions responsible for the bindings
of the various traits models can add additional Python attributes,
such as \pyLstinline{x}, \pyLstinline{y}, and
\pyLstinline{approximate}, which expose the member functions
\cppLstinline{x()}, \cppLstinline{y()}, and
\cppLstinline{approximate()}, to the Python object that exposes the
type  \cppLstinline{Point_2}.

The concept \cppLstinline{AosHorizontalSideTraits\_2} requires the
provision of the \cppLstinline{Compare\_x\_on\_boundary\_2}
functor\footnote{A functor is a class that has one or more
  \cppLstinline{operator()} members; thus, it acts as a function.}
that has two implementations of \cppLstinline{operator()} as follows;
one overload accepts two ends of open or unbounded curves (refer to
the manual for the precise encoding of a curve end) and another
overload accepts a point and an end of a curve that is open or
unbounded at that end. The former compares the $x$-coordinates of the
curves at their respective limits or ends and the latter compares the
$x$-coordinates of the point and the $x$-coordinates of the curve at
its respective limit or end.
%
The concept \cppLstinline{AosIdentifiedHorizontalTraits\_2} indirectly
refines the concept \cppLstinline{AosHorizontalSideTraits\_2}
mentioned above. (The former would be used to define a traits model
that handles arcs on a sphere with poles on the left and right sides;
Figure~\ref{fig:atc:sphericalBoundary} depicts a cluster of concepts
for models that handle arcs on a sphere with poles on the top and
bottom sides.) In addition to the two implementations of the
overloaded members \cppLstinline{operator()} of the functor
\cppLstinline{Compare\_x\_on\_boundary\_2} it requires the provision
of a third implementation that accepts two points and compares their
$x$ coordinates. Similar to the case described in the previous
paragraph, where the class object that handles the bindings for the
\cppLstinline{Point_2} type must be available in two separate
functions, also here the class object that handles the bindings for
the functor \cppLstinline{Compare\_x\_on\_boundary\_2} must be
available in both functions
\cppLstinline{export_AosHorizontalSideTraits\_2()} and
\cppLstinline{export_AosIdentifiedHorizontalTraits\_2()}. Saving the
class object in one function so that it can be reused latter on in the
other addresses this necessity.

% -----------------------------------------------------------------------------
\subsection{Type Annotation}
\label{ssec:concepts:annotation}
% -----------------------------------------------------------------------------
\cgal{} is a large library to start with. It is divided into
approximately~150 packages with countless function and class
templates. The vast number of function and type instances that can be
defined using these templates can be overwhelming. The development of
non critical code that is based on \cgal{} can be expedited using
Python and \cgal{} Python bindings.  Using code completion (which is
based on type annotations) offered by several Python IDEs can
accelerate the development process much further. Type annotation in
Python refers to annotations of Python functions, arguments, and
variables in a way that can be used by various tools. It is an
optional feature of Python that has been introduced in Version~3.5 and
enhanced in successive versions. With this feature implemented a type
checker, a tool separate from the Python interpreter, can be used to
statically analyze code base to find bugs during early development
stages and spelunk code of large projects. Naturally, it can be used
by IDEs to interactively provide hints and suggestions to programmers
and type check the code as it is being developed. Annotations can and
should be placed near the code if possible. Observe that the Python
interpreter ignores these annotations. Annotations can also be placed
in external files (which have the extension \lstinline{.pyi}),
referred to as stubs, for cases where the code base is inaccessible,
such as the case with bindings. Stub files contain the signatures of
the annotated functions with the body of the functions discarded. They
also list the attributes of annotated types. In our case they contain
the annotated interface of the \cgal{} bindings. Stub files have the
same syntax as regular Python modules. There is one feature of the
typing module that is different in stub files, namely, the
\pyLstinline{@overload} decorator. This decorator allows describing
functions that have different combinations of argument types.
Consider, for example, the overloaded free C++ functions
\cppLstinline{overlay()}; see Section~\ref{sec:code:overlay}. The
annotation for the corresponding Python function
\pyLstinline{overlay()} in the stub file is shown in
Listing~\ref{lst:py:overloading}.

\begin{lstfloat}
\begin{lstlisting}[style=pythonFramedStyle,basicstyle=\normalsize,%
                   label={lst:py:overloading},%
                   caption={Type annotation of the overloaded overlay function.}]
@overload
def overlay(r:Arrangement_2, b:Arrangement_2, o:Arrangement_2) -> None: pass
@overload
def overlay(r:Arrangement_2, b:Arrangement_2, o:Arrangement_2, t:Arr_overlay_traits) -> None: pass
@overload
def overlay(r:Arrangement_2, b:Arrangement_2, o:Arrangement_2, t:Arr_overlay_function_traits) -> None: pass
\end{lstlisting}
\end{lstfloat}

Annotating a Python type that wraps a C++ type, which is a model of
some concepts is guided by the concepts. Similar to the C++ code that
generates the bindings, we have introduced a tight coupling between
concepts and Python type annotations; see
Section~\ref{sec:code:concepts}. As opposed to the C++ binding code,
we have introduced a framework for the automatic generation of the
annotation stubs. The framework includes a Python script called
\bashLstinline{generate.py} that accepts as input static data files in
json format that describe the concepts, their refinement relations,
the exposed C++ classes, and their medeling relations with
concepts. The script generates annotation stubs for the bindings of
modules selected by the user. As mentioned in Section~\ref{sec:gen},
\cmake{} is used to generate the native build environment. We have
augmented the scope of \cmake, and now \cmake{} is also used to
generate the annotation stubs. Before we delve into the deatils of our
annotation stub generation system, we describe in the following two
paragraphs an unsuccessful attempt, which resembles the C++ binding
code. It reached a deadend because of limitations of Python.

Imagine that for each C++ concept we introduce a Python annotation
class that annotates all the requirements of the concept. The
annotated Python classes do not exist; these annotation classes are
introduced to facilitate the creation of the annotation classes of
real Python classes. Every annotation-class that annotates a certain
model type, say \cppLstinline{T}, inherits from all the
annotation-classes that correspond to the concepts that
\cppLstinline{T} models. For example, the type annotation for the
concept \cppLstinline{AosBasicTraits_2} contains the annotations the
nested type \cppLstinline{Point_2}, which in turn, contains
annotations for the default constructor and copy constructor of the
type \cppLstinline{Point_2} defined by every model of the concept
\cppLstinline{AosBasicTraits_2}; see
Listing~\ref{lst:py:aosBasicTraitsErr} for partial
annotations. Observe that the annotated Python class does not exist in
reality, thus the \pyLstinline{_} prefix in the name of the annotation
class \pyLstinline{_AosBasicTraits_2} in the listing. Assume that
bindings are generated for the
\cppLstinline{Arrangement_2<GeometryGtaits_2, Dcel>} type instantiated
with a geometry traits type that is an instance of the
\cppLstinline{Arr_segment_traits_2} class template.  This type models
several concepts; see, e.g., Figure~\ref{fig:atc:central}. Partial
erroneous Python type annotation for the type
\pyLstinline{Arr_segment_traits_2} is shown in
Listing~\ref{lst:py:aosSegmentTraitsErr}.

\begin{lstfloat}
  \begin{draft}
  \begin{lstlisting}[style=pythonFramedStyle,basicstyle=\scriptsize,%
                     %backgroundcolor=\color{red!20},
                     label={lst:py:aosBasicTraitsErr},%
                     caption={Erroneous partial type annotation for the attributes that correspond to the concept \cppLstinline{AosBasicTraits_2}.}]
class _AosBasicTraits_2():
  class Point_2():
    def __init__(): ...
    def __init__(const Point_2&): ...
  class Equal_2():
    @overload
    def __call__(self, p: _AosBasicTraits_2.Point_2, q: _AosBasicTraits_2.Point_2) -> bool: ...
    @overload
    def __call__(self, p: _AosBasicTraits_2.X_monotone_curve_2, q: _AosBasicTraits_2.X_monotone_curve_2) -> bool: ...
  def equal_2_object(self) -> _AosBasicTraits_2.Equal_2: ...
  \end{lstlisting}
  \end{draft}
\end{lstfloat}

\begin{lstfloat}
  \begin{draft}
  \begin{lstlisting}[style=pythonFramedStyle,basicstyle=\scriptsize,%
                     %backgroundcolor=\color{red!20},
                     label={lst:py:aosSegmentTraitsErr},%
                     caption={Erroneous partial Python type annotation for the type \pyLstinline{Arr_segment_traits_2}.}]
class Arr_segment_traits_2(AosBasicTraits_2._AosBasicTraits_2,
                          AosXMonotoneTraits_2._AosXMonotoneTraits_2,
                          AosTraits_2._AosTraits_2,
                          AosLandmarkTraits_2._AosLandmarkTraits_2,
                          AosVerticalSideTraits_2._AosVerticalSideTraits_2,
                          AosVerticalSideTraits_2._AosHorizontalSideTraits_2) :
  class Point_2(AosBasicTraits_2._AosBasicTraits_2.Point_2):
    def x(self) -> Ker.FT: ...
    def y(self) -> Ker.FT: ...
  \end{lstlisting}
  \end{draft}
\end{lstfloat}

Unfortunately, the code presented in
Listing~\ref{lst:py:aosBasicTraitsErr} and
Listing~\ref{lst:py:aosSegmentTraitsErr} does not work. Python and static
type checkers for Python (e.g., mypy)\footnote{See, e.g.,
  \url{https://mypy.readthedocs.io/en/stable/}.} have made significant
progress since Python was conceived. However, even with the
introduction of the \pyLstinline{Typing} module, and in particular the
\pyLstinline{Generic} class within,\footnote{See
  \url{https://docs.python.org/3/library/typing.html\#module-typing}.}
there are still barriers that cannot be bridged. The barrier that we
encountered is related to deep nesting. The culprit lies in Line~7 and
Line~9 of Listing~\ref{lst:py:aosBasicTraits}. In Line~7, for example,
the class \pyLstinline{Point_2} nested in the utility class
\pyLstinline{_AosBasicTraits_2} is refered to from a method of the
class \pyLstinline{Equal_2}, which is nested in
\pyLstinline{_AosBasicTraits_2}. An object of type
\pyLstinline{Arr_segment_traits_2.Equal_2} was not recognized
(probably because \pyLstinline{Arr_segment_traits_2.Equal_2} is not
substituted by \pyLstinline{_AosBasicTraits_2.Equal_2} even though
\pyLstinline{Arr_segment_traits_2} is derived from
\pyLstinline{_AosBasicTraits_2}.) and our ambitious attempt failed
flat.

Assume that two users select the \iiDArrangementsPackage{} module for
binding generation. In particular, one user chooses to generate
bindings for the \cppLstinline{Arrangement_2<GeometryGtaits_2, Dcel>}
type instantiated, where \cppLstinline{GeometryGtaits_2} is
substituted with an instance of the
\cppLstinline{Arr_segment_traits_2} class template and the other user
chooses to generate bindings for the
\cppLstinline{Arrangement_2<GeometryGtaits_2, Dcel>} type, where
\cppLstinline{GeometryGtaits_2} is substituted with an instance of the
\cppLstinline{Arr_Bezier_curve_traits_2} class template; see
Listing~\ref{lst:py:aosSegmentTraits} and
Listing~\ref{lst:py:aosBezierTraits} for the corresponding partial
valid Python type annotations generated by our system. The input data
files describe the concepts and the classes. More specifically, for
each concept $C$, their exists a record that lists the requirements of
the concept $C$ and the concepts the $C$ refines. For each type $T$,
there exists a record that lists the concepts that $T$ models and the
additional members and types nested in $T$ that are not covered by the
concepts. Type annotations are, naturally, generated only for the
types. An important feature of the system is the ability to merge
methods of a class that appears both as a nested type of another class
and as a type required by a concept. An example of such a scenario is
the \cppLstinline{Point_2} type required by the concept
\cppLstinline{Aosbasictraits_2} and nested in every geometry traits
class, e.g., \cppLstinline{Arr_segment_traits_2} and
\cppLstinline{Arr_Bezier_curve_2}.

\begin{lstfloat}
  \begin{lstlisting}[style=pythonFramedStyle,basicstyle=\scriptsize,%
                     %backgroundcolor=\color{red!20},
                     label={lst:py:aosSegmentTraits},%
                     caption={Valid Python type annotation for the type \pyLstinline{Arr_segment_traits_2}.}]
from typing import Iterator, overload
class Arr_segment_traits_2():
    class Point_2():
        @overload
        def __init__(self) -> None: ...
        @overload
        def __init__(self, p: Point_2) -> None: ...
        @overload
        def __init__(self, x: float, y: float) -> None: ...
        def x(self) -> FT: ...
        def y(self) -> FT: ...

    class Equal_2():
        def __call__(self, p: Point_2, Q: Point_2) -> Boolean: ...

    def equal_2_object(self) -> Equal_2: ...
  \end{lstlisting}
\end{lstfloat}

\begin{lstfloat}
  \begin{lstlisting}[style=pythonFramedStyle,basicstyle=\scriptsize,%
                     %backgroundcolor=\color{red!20},
                     label={lst:py:aosBezierTraits},%
                     caption={Valid Python type annotation for the type \pyLstinline{Arr_segment_traits_2}.}]
from typing import Iterator, overload
class Arr_Bezier_curve_traits_2():
    class Point_2():
        @overload
        def __init__(self) -> None: ...
        @overload
        def __init__(self, p: Point_2) -> None: ...
        def approximate(self) -> list[float]: ...

    class Equal_2():
        def __call__(self, p: Point_2, Q: Point_2) -> Boolean: ...

    def equal_2_object(self) -> Equal_2: ...
  \end{lstlisting}
\end{lstfloat}

While using a code-generation approach adds a level of complexity to
our binding system, it achieves the goal, it is flexible, and it
posses an additional advantage. Once the input data files for specific
concepts and models are in place, they can be used to automatically
generate the documentation for those models and concept; we certainly
plan to persue this goal in the future. The use of the type-annotation
subsystem, and in particular the \bashLstinline{generate.py} script or
the script that generates the documentation that we intend to develop
are not limited to \cgal{}---they can be used by any binding
generation system that generates Python bindings for generic C++ code.
