% =============================================================================
\section{Python bindings---comparison of tools}
\label{sec:comp}
% =============================================================================
There are several different tools and options for generating Python
byndings.  The following is by no means an exhaustive list of such
tools; however, it gives a taste of the rich possibilities one faces
when there is a need to make C++ code accessible to Python developers.
One goal that all these tools share is transforming Python/C API
into a relatively more user-friendly interface.
Most listed tools are designed to wrap C++ interfaces non-intrusively;
that is, the C++ code of an entity does not change when it is wrapped.

% -----------------------------------------------------------------------------
\subsection{Various Tools}
\label{ssec:comp:tools}
% -----------------------------------------------------------------------------
% ctypes
\pyLstinline{ctypes} is a foreign function module for Python that
provides C compatible data
types.\footnote{See \url{https://docs.python.org/3/library/ctypes.html\#module-ctypes}.}
It supports the loading of shared libraries and marshaling data
between Python and C.\footnote{Marshaling is the process of
transforming the memory representation of an object into a data format
suitable for storage or transmission; see
\url{https://en.wikipedia.org/wiki/Marshalling\_(computer\_science)}.}
The \pyLstinline{ctypes} module is part of the Python standard
library; thus, it can be used to wrap shared libraries in pure Python.
Using the \pyLstinline{ctypes} module requires manual programming
labour and detailed knowledge of the \pyLstinline{ctypes}
interface. However, it does not require writing C code (in addition to
existing C code that needs to be wrapped).

% CFFI
\pyLstinline{cffi} is another Python module that generates Python
bindings (it stands for C Foreign Function Interface). It interacts
with almost any C code from Python, based on C-like declarations that
can often be copy-pasted from header files. It is compatible with both
PyPy and CPython; Moreover, in the PyPy (see~\ref{ssec:intro:bind})
reference implementation it is a first class citizen (and provided by
defaut), thus, exhibts outstanding preformance. Compared
to \pyLstinline{ctypes}, the \pyLstinline{cffi} module takes a more
automated approach to generate Python bindings and it minimized the
extra bits of API that needs to be mastered; thus, it scales better to
larger projects than \pyLstinline{ctypes}. Like \pyLstinline{ctypes},
using \pyLstinline{cffi} directly interfaced with C libraries
only. Warapping C++ libraries with any one of these modules requires
the provision of another layer of C wrapper around the C++ code.

% cppyy
\pyLstinline{cppyy} is yet another external module for generating
bindings.\footnote{See \url{https://cppyy.readthedocs.io}.} It is an
automatic, run-time, C++ Python bindings generator, for calling C++
from Python and Python from
C++.\footnote{See \url{https://cppyy.readthedocs.io}.} Run-time
generation enables detailed specialization for higher performance,
lazy loading for reduced memory use in large scale projects,
Python-side cross-inheritance and callbacks for working with C++
frameworks, run-time template instantiation, automatic object
downcasting, exception mapping, and interactive exploration of C++
libraries. \pyLstinline{cppyy} delivers this without any language
extensions, intermediate languages, or the need for boiler-plate
hand-written code. For design and performance, see~\cite{ld-hppcn-16},
albeit that the \pyLstinline{cppyy} performance has been vastly
improved since.  Like \pyLstinline{cffi}, \pyLstinline{cppyy} was
designed to ease rge binding development on Python programmer,
minimizing the need for extra C++ code; it is also compatible with
both CPython and PyPy.

% Cython
A completely different approach is adopted by
Cython~\cite{bbcds-cbbw-11}, which refers to a programming language
and to an optimising static compiler for the Cython
language.\footnote{See \url{https://cython.org/}.}  The Cython
language is a superset of the Python language. It extends Python with
explicit type declarations of native C/C++ types.  Cython is a
compiled language used to generate CPython extension modules.  The
Cython compiler generates C/C++ code from Cython code, which in turn
compiles with any C/C++ compiler, and automatically wrapped in
interface code, producing extension modules that can be loaded and
used by regular Python code using the \pyLstinline{import} statement,
but with significantly less computational overhead at run time. Cython
also facilitates wrapping independent C or C++ code into
python-importable modules.  With Cython it is possible to call back
and forth between Python code, Cython code, and native library code
origianlly developed in Fortrant, C, or C++.

% SIP
SIP is a collection of tools that makes it easy to create Python
bindings for C and C++ libraries.\footnote{See
  \url{https://pypi.org/project/sip/}.}  It was originally developed
in 1998 to create PyQt, the Python bindings for the Qt toolkit, but
can be used to create bindings for any C or C++ library. SIP comprises
a set of build tools and a module called sip. The build tools process
a set of specification files and generates C or C++ code, which is then
compiled to create the bindings extension module. Several extension
modules may be installed in the same Python package. Extension modules
can be built so that they are are independent of the version of Python
being used.
% In other words a wheel created from them can be installed with any
% version of Python starting with v3.5.

% PyBindGen

% SWIG
SWIG is a software development tool that connects programs written in
C and C++ with a variety of high-level programming
languages.\footnote{See \url{http://www.swig.org/}.}  It is defferent
than any of the others tools listed here.  SWIG is used to create
bindings to C and C++ code for different types of target languages,
and not only Python, including common scripting languages such as
Javascript, Perl, PHP, Python, Tcl, or Ruby. The list of supported
languages also includes non-scripting languages such as C\#, D, Go
language, Java including Android, Lua, OCaml, Octave, Scilab and
R. Also several interpreted and compiled Scheme implementations
(Guile, MzScheme/Racket) are supported. SWIG is most commonly used to
create high-level interpreted or compiled programming environments,
user interfaces, and as a tool for testing and prototyping C/C++
software. SWIG is typically used to parse C/C++ interfaces and
generate the 'glue code' required for the above target languages to
call into the C/C++ code. SWIG can also export its parse tree in the
form of XML. SWIG is free software and the code that SWIG generates is
compatible with both commercial and non-commercial projects.

% Boost.Python
Boost is a large and complex suite of utility libraries that works
with almost every C++ compiler in existence. Boost.Python~\cite{bhsbp-ag-03} is a member of the Boost suit that enables seamless interoperability between C++
and the Python programming language.\footnote{See
  \url{https://www.boost.org/doc/libs/1_70_0/libs/python/doc/html/index.html}.}
similar to \pyLstinline{cppyy}, Boost.Python focuses at C++; unlike
\pyLstinline{cppyy} it uses C++ to specify and build bindings, taking
advantage of the metaprogramming tools and polymorphism in C++.

The library includes support for:
\begin{itemize}
\item References and Pointers
\item Globally Registered Type Coercions
\item Automatic Cross-Module Type Conversions
\item Efficient Function Overloading
\item C++ to Python Exception Translation
\item Default Arguments
\item Keyword Arguments
\item Manipulating Python objects in C++
\item Exporting C++ Iterators as Python Iterators
\item Documentation Strings
\end{itemize}

% PyBind11
pybind11 is a lightweight header-only library that exposes C++
types in Python and vice versa, mainly to create Python bindings of
existing C++
code.\footnote{See \url{https://github.com/pybind/pybind11}.} Its
goal, like the goal of Boost.Python, is to minimize boilerplate code
in traditional extension modules by inferring type information using
compile-time introspection. Its syntax is als similar to the syntax of
Boost.Python. In fact, it is a tiny self-contained version of
Boost.Python with everything stripped away that isn't relevant for
binding generation. Without comments, the core header files only
require approximately 4K lines of code and depend on Python and the
C++ standard library. This compact implementation was possible thanks
to some of the new C++11 language features (specifically: tuples,
lambda functions and variadic templates). Since its creation, this
library has grown beyond Boost.Python in several ways, leading to
simpler binding code in many common situations.

We have concentrated at three tools, namely (i) SWIG, (ii) PyBind11,
and (iii) Boost.Python and conducted several experiments to compare
the performance of the respective generated bindings; see
Section~\ref{ssec:comp:exp}.

% -----------------------------------------------------------------------------
\subsection{Experiments}
\label{ssec:comp:exp}
% -----------------------------------------------------------------------------
C++ Python bindings for a certain subset of \cgal{} packages are
available as part of an experimental project that uses SWIG. While the
project has been conceived more than a decade ago, it contains
bindings only for a limited subset of \cgal{} packages. Naturally,
embarking on the existing venture might have saved time in developing
new bindings, on the one hand, and in exploiting the new bindings by
our users on the other. Another advantage of using SWIG, and probably
the reason for choosing SWIG in the first place in the aforementioned
project, is the support for multiple languages. Once bindings have
been developed for one language, generating bindings for additional
languages does not require additional development resources. However,
as our experiments show, this profiteering has its cost---the
generated bindings are not as time or space efficient as their
counterparts generated with other tools. The PyBind11 tool is
relatively new. Both the PyBind11 and the Boost.Python tools are
widely spread and tested, and actively maintained and enhanced. They
are developed using advanced features of C++11 and higher
versions. These two tools share the same interface; that is, the C++
code used to generate the bindings using both tools is similar and in
some cases even identical. Also, switching between using bindings
generated by PyBind11 and bindings generated by Boost.Python is
seemless---it amounts to replacing one libarary by another. As we are
very much familiar with developing code that exploits advanced feature
of C++, developing bindings using these tools does not present a major
hurdle for us. One the other hand, developing bindings using SWIG,
requires writing code that describe the bindings in a propriatary
language.  Some developers may avoid using Boost.Python, because it
requires dealing with a large and complex system. This does not
present a drawback for us, because \cgal{} itself depends on several
other libraries of Boost; thus, we are familiar with Boost and must
deal with it anyway.  The compatibility of Boost with almost every C++
compiler in general has its cost; arcane template tricks and
workarounds are necessary to support the oldest and buggiest of
compiler specimens.  Now that C++11-compatible compilers are widely
available, this heavy machinery has become an excessively large and
unnecessary dependency. However, our main concern was the efficiency
of the generated bindings in terms of execution time and memory space.

We have conducted two sets of experiments. In the first set we
compared the time it took to refer to C++ objects from Python code and
in the second set we compared the time it took to call C++ functions
from Python code. The first set was further divided into two subsets.
In the first subset we compared the time it took to make many
references to many objects and in the second subset we compared the
time it took to make many references to a single object.

\begin{table}
  \caption{References}
  \centering\begin{tabular}{|l|r|r|r|}
    \hline
    & \textbf{pybind11} & \textbf{boost} & \textbf{SWIG}\\
    \hline
    \hline
    Many references to many objects    & 6.03125  & 4.453125 & 15.6875\\
    Many references to a single object & 4.90625  & 2.796875 &  9.640625\\
    \hline
  \end{tabular}
\end{table}

The second set was further divided into too subsets as well. In the
first subset we compared the time it took to call many small
functions, and in the second subset we compared the time it took to
call few large functions. We also used the second set of experiments to
measure the time it took to compute the various tasks in native Python.

\begin{table}
  \caption{Function calls}
  \centering\begin{tabular}{|l|r|r|r|r|}
    \hline
    & \textbf{pybind11} & \textbf{boost} & \textbf{SWIG} & \textbf{Native}\\
    \hline
    \hline
    Many small tasks & 26.71875 & 13.65625 & 13.140625 & 13.359375\\
    One large task   & 0.71875  & 0.71875  & 0.734375  & 17.84375\\
    \hline
  \end{tabular}
\end{table}

% Boost.Python handles shared\_ptr and other smart pointers extremely gracefully
