\begin{abstract}
  We introduce bindings that enable the convenient, efficient, and
  reliable use of software modules of \cgal{} (Computational Geometry
  Algorithm Library), which are written in C++, from within code written in
  Python.
  %
  There are different tools that facilitate the creation of such
  bindings. We present a short study that compares three main tools,
  which leads to the tool of choice.
  % The implementation of practical algorithms and data structures in
  Computational geometry presents tremendous difficulties, such as
  obtaining stable software despite the use of (inexact) floating
  point arithmetic, found in standard hardware, and obtaining complete
  software that handles all degenerate cases, which typically are in
  abundance. The code of \cgal{} extensively uses function and class
  templates in order to handle these difficulties, which implies that
  the programmer has to make many choices that are resolved during
  compile time (of the C++ modules). While bindings take effect at run
  time (of the Python code), the type of the C++ objects that are
  bound, must be known when the bindings are generated, that is, when
  they are compiled. The types of the bound objects are instances
  (instantiated types) of C++ function and class templates. The number
  of object types that potentially can be bound is enormous; thus, the
  generation of the bindings in advance is practically impossible. We
  present a system that dynamically generates bindings for desired
  object types according to user prescriptions, which enables the
  convenient use of any subset of bound object types concurrently.
\end{abstract}
