% =============================================================================
\section{Binding Code}
\label{sec:bind-code}
% =============================================================================
Boost.Python allows users to expose C++ classes and functions to
Python using nothing more than a C++ compiler. It exploits
metaprogramming techniques to implement a rich set of features and
high-level user-friendly interface. Boost.Python extracts as much as
information as possible from the source code to be wrapped. This
approach is referred to as \emph{user guided wrapping}. Additional
information that cannot be automatically extracted is explicitly
supplied by the user. The interface specification is written in the
same full-featured C++ language as the code being exposed, which has
the potential of generating efficient binding. In this section we
describe some of the techniques we use to ensure that.

% -----------------------------------------------------------------------------
\subsection{Return value policy}
\label{sec:bind-code:rvp}
% -----------------------------------------------------------------------------
Python and C++ use fundamentally different ways of managing the memory
and lifetime of objects. This can lead to issues when creating
bindings for functions that return a non-trivial type. Just by looking
at the type, it is unclear whether Python should take charge of the
returned value and eventually free its resources, or whether this is
handled by the C++ code. When writing code in C++, it is usually
considered a good practice to use smart pointers, which exactly
describe ownership semantics. Still, even good C++ interfaces use raw
references and pointers sometimes. In some cases, in order to assure
proper memory management of a return value from a function, explicitly
specifying a return value policy is needed. The policy may depend on
whether the returned object is a newly created object, a reference to
some already existing (internal) object, or a const reference to some
existing object. Additionally, in some cases we need to tie the
lifetime of the result to the lifetime of the arguments, or the other
way around.

For some functions of kernel objects the kind of return value depends
on the kernel type; thus, the return value policy needs to be set
accordingly. For example, when the kernel type is
\cppLstinline{Exact\_predicates\_inexact\_constructions\_kernel}, the
return type of some functions that return a coordinate or a
coefficient is a const reference, while when the kernel type is
\cppLstinline{Exact\_predicates\_exact\_constructions\_kernel} the same
functions return an object by value.

This is why for these functions a
\myLstinline{Kernel_return_value_policy} type is being passed as the
return value policy. The type is defined differently (via preprocessor
macros) based on the kernel being exposed.

% -----------------------------------------------------------------------------
\subsection{Scopes}
\label{sec:bind-code:scopes}
% -----------------------------------------------------------------------------
The Boost.Python mechanism that depends on \cppLstinline{bp::scope} is
used to (i) reflect nesting in C++ code that uses \cgal, and (ii)
prevent name conflicts between bound entities in distinct modules; see
Section~\ref{sec:bind-gen:scopes}. The latter requires a slight
intervention on our part as explained next. The code that binds all
functions and classes of different modules is nested in distinct C++
forced scopes. Within a (C++) scope we first set the Python namespace
to match the name of a module. Then, we bind all functions and classes
of \cgal{} packages associated with the module. The following code
except shows how it is done for the \iiDandiiiDGeometryKernelPackage{}
module:

\begin{lstlisting}[style=cppNumberedStyle]
{
  SET_SCOPE("Ker")
  export_kernel();
#ifdef CGALPY_KERNEL_INTERSECTION_BINDINGS
  export_kernel_intersections_2();
  export_kernel_intersections_3();
#endif
}
\end{lstlisting}

For every module there exsists a function called
\cppLstinline{export_<name>}, where \cppLstinline{<name>} is the
module. This function contains the binding code for the essential
functions and classes of the module. Recall that the
\iiDandiiiDGeometryKernelPackage{} module is associated with the
\myLstinline{KERNEL\_INTERSECTION\_BINDINGS} flag, which determines
whether to generate bindings for intersections; hence, the
compile-time conditional calls to the functions
\cppLstinline{export_kernel_intersections_2()} and
\cppLstinline{export_kernel_intersections_3()}, which contains the
binding code for intersections.

%% All functions and classes related to some package were exposed inside
%% a \myLstinline{export_[package]()} function. All such calls are
%% happening in \myLstinline{export_module.cpp}. We wrap each call with a
%% local (c++) scope and call the \myLstinline{SET_SCOPE(module_name)}
%% macro, which causes every exposed function/class inside this local
%% scope to be exposed inside the \myLstinline{module_name} submodule of
%% the CGALPY module. This prevents name clashes of classes with the same
%% name in different packages.

% -----------------------------------------------------------------------------
\subsection{Avoiding Code Duplication}
\label{sec:bind-code:templates}
% -----------------------------------------------------------------------------
Some types have the exact same nested types and member functions
exposed, e.g., \cppLstinline{Triangulation\_2} and
\cppLstinline{Delaunay\_triangulation\_2} class templates each have
the method \cppLstinline{dimension()} exposed as the Python attribute
\pyLstinline{dimension}. In these cases we use a function template to
avoid code duplication. Following the example above, we have the
function \cppLstinline{export\_triangulation\_2<Triangulation>()},
where the template parameter \cppLstinline{Triangulation} is
substituted with one of the two-dimensional triangulation types
supported; see Table~\ref{tab:tri2:names}. This function template
accepts as its single parameter the binding class object, which is
constructed separately.

\begin{lstlisting}[style=cppNumberedStyle]
typedef CGAL::Triangulation_2<Kernel>              Triangulation_2;
typedef CGAL::Delaunay_triangulation_2<Kernel>     Delaunay_triangulation_2;

auto c0 = bp::class_<Triangulation_2>("Triangulation_2");
auto c1 = bp::class_<Delaunay_triangulation_2,
                     bp::bases<Triangulation_2>>("Delaunay_triangulation_2");

export_triangulation_2<Triangulation_2>(c0);
export_triangulation_2<Delaunay_triangulation_2>(c1);
\end{lstlisting}

% -----------------------------------------------------------------------------
\subsection{Substitution Failure Is Not an Error}
\label{sec:bind-code:sfinae}
% -----------------------------------------------------------------------------
Substitution failure is not an error (SFINAE) refers to a situation in
C++ where an invalid substitution of template parameters is not in
itself an error. In particular, it applies during a process called
\emph{overload resolution}.  In order to compile a function call, the
compiler creates a set of candidate functions. This set includes all
function implementations the name of which match the call and that can
be accessed by the caller. Then the compiler selects the most viable
one according to some criteria. When substituting the explicitly
specified or deduced type for the template parameter fails, the
specialization is discarded from the overload set instead of causing a
compile error.\footnote{For more information see, e.g.,
  \url{https://en.cppreference.com/w/cpp/language/sfinae}.}

The function \cppLstinline{CGAL::do_intersect(o1, o2)} is overloaded
with several implementations that handle different combination of
types of arguments; however, not every combination of two types is
implemented. For a complete list of valid combinations refer to the
reference manual at
\url{https://doc.cgal.org/latest/Kernel_23/group__do__intersect__linear__grp.html}. We
attempt to create bindings for all combinations and avoid getting
compilation errors for unsupported combinations with the help of
SFINAE as follows. Consider two types \cppLstinline{Type1} and
\cppLstinline{Type2}, and assume that the overloaded version
\cppLstinline{CGAL::do_intersect(Type1 o1, Type2 o2)} is not
implmented but all other three are implemented.

\begin{lstlisting}[style=cppNumberedStyle]
template <typename T1, typename T2>
void bind_do_intersect_2T(decltype(CGAL::do_intersect<Kernel>(T1(), T2()))) {
  bp.def<bool(const T1&, const T2&)>("do_intersect", &CGAL::do_intersect<Kernel>);
}

template <typename, typename>
void bind_do_intersect_2T(...) {}

template <typename T1>
void bind_do_intersect_1T() {
  bind_do_intersect_2T<T1, Type1>(true);
  bind_do_intersect_2T<T1, Type2>(true);
}

void bind_do_intersect() {
  bind_do_intersect_1T<Type1>();
  bind_do_intersect_1T<Type2>();
}
\end{lstlisting}

First the function \cppLstinline{bind_do_intersect()} calls the
instance of \cppLstinline{bind_do_intersect_1T<T1>()}, where
\cppLstinline{T1} is substituted with \cppLstinline{Type1}.  Then, the
latter makes two calls and passs a \cppLstinline{bool} argument of
value \cppLstinline{true}. First it calls the instance of
\cppLstinline{bind_do_intersect_2T<Type1, T2>(true)}, where
\cppLstinline{T2} is also substituted with \cppLstinline{Type1}; then,
it calls the instance of \cppLstinline{bind_do_intersect_2T<Type1,
  T2>(true)}, where \cppLstinline{T2} is substituted with
\cppLstinline{Type2}. The function
\cppLstinline{bind_do_intersect_2T()} accepts a parameter of type
\cppLstinline{bool} and is overloaded with two implementations.  The
second uses \emph{variadic arguments} and serves as a fall-through
when the evaluation of the first implementation fails during the
resolution process.\footnote{For more information on \emph{variadic
    arguments} see
  \url{https://en.cppreference.com/w/cpp/language/variadic_arguments}.}
The paramater type of the first implementation is defined as the
return type of \cppLstinline{CGAL::do_intersect<Kernel>(T1(),
  T2())}. When \cppLstinline{T1} and \cppLstinline{T2} are substituted
with \cppLstinline{Type1} and \cppLstinline{Type2}, respectively, the
evaluation fails, and the blank implementation is selected and
invoked.  In all other three cases the evaluation succeeds. As
variadic parameters have the lowest rank for the purpose of overload
resolution, the first implementation will alsways be selected for
valid combinations.

%% templated with some supported type
%% {T1}. \myLstinline{bind_do_intersect_1T()}, templated with
%% \myLstinline{T1}, calls \myLstinline{bind_do_intersect_2T} templated
%% with \myLstinline{T1} and some supported type \myLstinline{T2}, and a
%% \myLstinline{bool} argument.  The declaration of
%% \myLstinline{bind_do_intersect_2T}, templated with both
%% \myLstinline{T1} and \myLstinline{T2} is only valid when templated
%% with a combination of types for which an overloaded implementation of
%% \myLstinline{CGAL::do_intersect} exists. Otherwise, due to SFINAE, the
%% blank implementation template of \myLstinline{bind_do_intersect_2T}
%% expecting (...) as an argument will be invoked (it would normally not
%% be invoked as variadic parameters have the lowest rank for the purpose
%% of overload resolution).

% -----------------------------------------------------------------------------
\subsection{Arrangement Extension}
\label{sec:bind-code:ae}
% -----------------------------------------------------------------------------
In many applications it is necessary to extend the type that
represents certain \cgal{} data structures with some properties. In
\cgal{} it is more convenient, and thus more common, extending the
types that represent the topological features of the data structures
(rather then the types that represent the geometric features), for
example, the DCEL of the arrangement or the triangulation data
structure of a triangulation. When developing code in pure C++, there
is no restriction on the property used to extend the data structure.
However, when the data structure interface is wrapped with Python
bindings, the property type must be known when the bindings are
compiled. Nevertheless, we support extentions with generic Python
objects. If a certain feature must be extended with a certain
property, the type of which is defined in the C++ code, it is always
possible to use an external propery map, where the properties are
indexed by a Python object, such as an integer or a string. When
generating the bindings the user must set the flag that enables the
extension (unless extension is the default setting). Attaching indices
to and extracting an index from a feature is done via the
\cppLstinline{set_data()} and \cppLstinline{data()} exposed functions.

Consider, for example, the \cppLstinline{Arrangement\_2<Traits,Dcel>};
see Section~\ref{ssec:modules:aos2}. An instance of this template is
used to represent an arrangement embedded in the plane. The following
example ...

- Extending the 2d and 3d vertex and face triangulation types => change the title to be more general

% -----------------------------------------------------------------------------
\subsection{Arrangement Overlay Traits}
\label{sec:bind-code:overlay}
% -----------------------------------------------------------------------------
Formally, the map overlay of two spatial subdivisions $\calS_1$ and
$\calS_2$ is a third subdivision $S$, such that there is a cell $c$ in
$\calS$ iff there are cells $c_1$ and $c_2$ in $\calS_1$ and
$\calS_2$, respectively, such that $c$ is a maximal connected
component of $c_1 \cap c_2$. Arrangements are subdivisions of some
ambiant space (see Section~\ref{ssec:modules:aos2}) and computing the
overlay of two arrangements is useful for many applications, including
the application of Boolean operations; see
Section~\ref{ssec:modules:bso2}. Indeed, the \iiDArrangementsPackage{}
\cgal{} package supports an operation that computes the map overlay of
two arrangements in form of the free function template
\cppLstinline{CGAL::overlay()}. In particular, the call
\cppLstinline{CGAL::overlay(arr_r, arr_b, arr_o, traits)} constructs
the arrangement \cppLstinline{arr_o}, which is the overlay of two
input arrangement \cppLstinline{arr_r} and \cppLstinline{arr_b}. The
last argument, namely \cppLstinline{traits}, enables the population of
properies in the extended DCEL records of the output arrangement from
properties of the DCEL records of the input arrangements. All three
arrangements must use the same geometric primitives.  In other words,
their types are instances of the \cppLstinline{Arrangement_2<Traits,
  Dcel>} class template, where the \cppLstinline{Traits} parameter is
substituted with three geometry-traits classes, respectively. The
geometry-traits classes of the input arrangements must be convertible
to the geometry-traits class of the resulting arrangement. This is the
only restriction that must be enforced when the function is used in
pure C++ code.  We only wrap instances of the overlay function with
Python bindings, where all three arrangements use the same
geometry-traits classes.

The type of the overlay traits models the \cppLstinline{OverlayTraits}
concept, which requires the provision of ten functions that handle all
possible cases as listed below. Let $v_r$, $e_r$, and $f_r$ denote
input \textcolor{red}{red} features, i.e.,~a vertex, an edge, and a
face, respectively, $v_b$, $e_b$, and $f_b$ denote input
\textcolor{blue}{blue} features, and $v$, $e$, and $f$ denote output
features.
\begin{compactenum}
\item A new vertex $v$ is induced by coinciding vertices $v_r$ and $v_b$.
\item A new vertex $v$ is induced by a vertex $v_r$ that lies on an edge $e_b$.
\item An analogous case of a vertex $v_b$ that lies on an edge $e_r$.
\item A new vertex $v$ is induced by a vertex $v_r$ that is contained in a
  face $f_b$.
\item An analogous case of a vertex $v_b$ contained in a face $f_r$.
\item A new vertex $v$ is induced by the intersection of two edges $e_r$ and
  $e_b$.
\item A new edge $e$ is induced by the (possibly partial) overlap of two edges $e_r$ and $e_b$.
\item A new edge $e$ is induced by the an edge $e_r$ that is contained in a
  face $f_b$.
\item An analogous case of an edge $e_b$ contained in a face $f_r$.
\item A new face $f$ is induced by the overlap of two faces $f_r$ and $f_b$.
\end{compactenum}
Evidently a custom C++ overlay traits cannot be defined in Python;
thus, the traits type must be known when the bindings are compiled. We
introduce a new class, namely \myLstinline{Arr_python_overlay_traits},
that is a model of the \cppLstinline{OverlayTraits} concept. Its
constructor accepts ten Python functions, each accepting two Python
objects and returning a Python object. As a result, the user (that is,
a Python programmer) can still computes the properties of the output
extended arrangement. Recall that these properties are represented by
Python objects to start with; see Section~\ref{sec:bind-code:ae}.

%% (the extended arrangement data type) Those 10 functions are saved as
%% class members and are then called in the respective 10 [overloaded]
%% versions of \myLstinline{create_vertex}, \myLstinline{create_edge},
%% \myLstinline{create_face object}) (the matching is done by the order
%% of the passed functions and the order of the [overloaded] versions in
%% the official concept reference).

%% This way the user can still define the way the result face/edge/vertex
%% data will be calculated.

% -----------------------------------------------------------------------------
\subsection{kd tree}
\label{sec:bind-code:kdtree}
% -----------------------------------------------------------------------------
The k-d tree dimension and the dimension of related classes is
templated and can thus only be decided at compile time via the cmake
variable \myLstinline{CGALPY_SPATIAL_SEARCHING_DIMENSION}.

In order to assert the dimension the bindings were compiled with, a
free function \myLstinline{get_spatial_searching_dimension()} was
exposed that returns the dimension as an \myLstinline{int}.

% -----------------------------------------------------------------------------
\subsection{\myLstinline{Python_distance}}
\label{sec:bind-code:distance}
% -----------------------------------------------------------------------------
Similarly to overlay traits, a user may want to define a custom
distance class to use with search queries. Here too, as defining a new
c++ class is not possible from python, we provide a similar solution:
A \myLstinline{Python_distance} class is exposed, which is a model of
\myLstinline{GeneralDistance} concept.
\url{https://doc.cgal.org/5.2/Spatial_searching/classGeneralDistance.html}
It expects 5 functions in its constructor, and those are saved as
class members and are used for the respective 5 operations [the
optional are not included] of the concept.

Making some things more pythonic:

% -----------------------------------------------------------------------------
\subsection{Iterators and circulators}
\label{sec:bind-code:iters}
% -----------------------------------------------------------------------------
For binding iterators we used the \myLstinline{bind_iterator}
templated function and its variations, in order to allow pythonic
iteration. Circulators too are artificially converted to iterators
supporting \myLstinline{__next__} and throwing a stop iteration when
reaching the first item again, by being wrapped in a templated
"\myLstinline{Iterator_from_circulator}" object which is the one being
exposed to python.

% -----------------------------------------------------------------------------
\subsection{Functions populating an output iterator}
\label{sec:bind-code:fnc-out}
% -----------------------------------------------------------------------------
Instead, a wrapper that expects a python list and populates it with
the results is being exposed.

% -----------------------------------------------------------------------------
\subsection{Functions expecting begin and end iterators}
\label{sec:bind-code:fnc-range}
% -----------------------------------------------------------------------------
Instead, a wrapper that expects a python list as an argument and calls
the function with the list's begin and end iterators is exposed.
