% =============================================================================
\section{Introduction}
\label{sec:intro}
% =============================================================================
Bindings are essentially wrapper libraries that bridge two
programming languages, so that a software module developed in one
language can be used in code written in another language, exploiting
(i) the strengths of both languages, and (ii) the availability of
modules developed in the former language, if such exist. C++ Python
bindings enable the invocation of C++ functions from Python code
taking advantage of Python's quick development cycles and C++'s high
performance.
%
% =============================================================================
\subsection{C++ Python Bindings}
\label{ssec:intro:bind}
% =============================================================================
Python, like many other languages, was originally defined by a
reference implementation. The original reference implementation of the
Python programming language is CPython. Todate, CPython is still the
default and commonly used reference implementation of the language,
although new rivals are emerging, see, e.g., PyPy.\footnote{PyPy is a
  replacement for CPython; see \url{https://www.pypy.org}.}. As the
Python language evolved its definition has tightened and a well
defined specification has materialized.

Python is sufficiently efficient for many tasks; nevertheless
low-level code written in Python tends to be too slow, largely due to
Python's extremely dynamic nature. In particular, low-level
computational loops are simply infeasible. In additional, the
magnitude of existing and well-tested code in Fortract, C, and C++, is
enourmous. The Python/C API, the module that enables using extension
modules, developed in Fortrant, C, or C++, from Python code, or
embedding Python code in code written in other languages is very rich.
Indeed, the Python/C API module of CPython exposes a vast amount of
details of CPython internals. It enables the efficient exploitation of
existing code in Fortrant, C, or C++, as well as the replacement of
critical sections written in these languages when speed is essential.

Wrapping existing code has traditionally been the domain of Python
experts due to the steep learning curve, which characterizes its
Python/C API. Although using such wrappers is possible without ever
knowing their internals, providing such wrappers creates a sharp line
between developers using Python and developers using C or C++ with
the Python/C API. Several tools that facilitate the creation of such
wrappers have been introduced; A sample of such tools are listed In
Section~\ref{sec:comp}.

% =============================================================================
\subsection{\cmake}
\label{ssec:intro:cmake}
% =============================================================================
\cmake{}~\cite{mh-mc-08} is an open-source, cross-platform suite of
tools designed to build, test, and package software. The suite of
\cmake{} tools were created to address the need for cross-platform
build environments for open-source projects. However, since it was
conceived over two decades ago, it has grown and become powerful,
rich, and flexible. It is now considered a modern tool; it is widely
spread and used by major commercial software products for their build
and test environments.

\cmake{}, for the most part, is a cross-platform build system
generator.  It is used to control the software build process using
simple platform and compiler independent configuration files, and it
generates native build environments of the user choice. Users build a
project by using \cmake{} to generate a build system for native tools
on their platforms. \cmake{} supports an interpreted, functional,
scripting language. The build process of a project is specified with
platform-independent \cmake{} text files, named
\myLstinline{CMakeLists.txt}, written in the \cmake{} language,
included in each directory of the source tree of the project. For
example, once the \cmake{} configuration files are in place for a
certain project, \cmake{} can be used to generate
\myLstinline{Makefile} files required by the \myLstinline{make}
utility of the GNU suite to compile a project on a Unix system using
\myLstinline{g++}. The command-line interface of the cross-platform
build system generator of \cmake{} is an executable called
\myLstinline{cmake}. It has a counterpart called
\myLstinline{cmake-gui}, which also includes a graphical user inteface
(GUI). To generate native project build files, the user invokes
\myLstinline{cmake} in a terminal and specifies the directory that
contains the root \myLstinline{CMakeLists.txt} file.

A key feature of \cmake{} is the ability to (optionally) place
compiler outputs (such as object files) outside the source tree. This
enables multiple and concurent builds from the same source tree and
cross-compilation. It also keeps the source tree clean and ensures
that removing a build directory will not remove the source files.  We
exploit this feature, and the scripting capabilities to conveniently
generate the C++ Python bindings for several \cgal{} objects
concurrently

% =============================================================================
\subsection{\cgal}
\label{ssec:intro:cgal}
% =============================================================================
\cgal{} (Computational Geometry Algorithm Library) is an open source
software project that provides easy access to efficient and reliable
implementations of geometric algorithms in the form of a C++
library~\cite{cgal:eb-21}. \cgal{} has evolved through the years and
now represents the state of the art in applied computational geometry
software. The software modules of \cgal{} rigorously adhere to the
generic programming paradigm to overcome problems encountered when
effective computational geometry software is implemented; they are
stable despite the use of (inexact) floating point arithmetic, found
in standard hardware, and they are complete as they handle all
degenerate cases, which typically are in abundance.

When running Python code that use bindings for some C++ modules, one
or more libraries that provide the bindings must be accessible.  A
software module that adheres to the generic programming paradigm
consists of function and class templates; these templates are
instantiated at compile time of the binding libraries. In other words,
the type of the C++ objects that are bound, that is, instances
(instantiated types) of C++ function and class templates, must be
known when the bindings are generated.

Generic programming identifies an abstraction that consists of a
formal hierarchy of polymorphic abstract requirements on data types,
referred to as \emph{concepts}, and a set of classes that conform
precisely to the specified requirements, referred to as
\emph{models}~\cite{a-gps-99}. When a class template is instantiated,
each one of its template parameters is substituted with a model of a
concept associated with the template parameter.

Instantiated types in \cgal{} are characterized by long instantiation
chains of C++ class templates. An instantiated type is a template that
has one or more template parameters and every parameter is substituted
by another type that is typically an instance of another template.
%
While the number of models of most concepts is small, the number of
potential types of objects that could be bound, and thus must be
supported by the bindings, is enormous. Offering bindings for all
these types in advance is practiaclly impossible. Moreover, in some
cases models need to be extended with types provided by the
user. Naturally, bindings for user-extended models must be generated
dynamically.

\cgal{} uses \cmake{} for the build process.  Since version~5.0,
\cgal{} is header-only by default, which means that an application
that depends on \cgal{} only depends on (i) the source code of \cgal{}
that resides in header file (which typically end with the suffix
\myLstinline{.hpp}), and (ii) native configuration files generated by
\cmake{} (in particular, the execution of \myLstinline{cmake}). In
other words, there is no need to compile any library before an
application that depends on \cgal{} and written in C++ is
built.\footnote{Some dependencies of \cgal{} might need to be
  installed in advanced}. On the other hand, when running an
application written in Python that depends on \cgal{} the libraries
that contain the bindings must be accessible. Typically, those
bindings are generated ahead of time with little knowledge about the
application itself. This would impose a severe limitation in cases
like ours, where the number of objects to be bound is large. In cases
where the type pf the objects to be bound are closely tied with the
application itself, generating bindings ahead of time is, naturally,
impossible. Our system enables easy and convenient (re)generation of
bindings, thus alleviating the burdon caused by these problems.

% =============================================================================
\subsection{Results}
\label{ssec:intro:results}
% =============================================================================
We introduce C++ Python bindings that enable the convenient,
efficient, and reliable use of \cgal{} software modules in Python's
code. There are different tools that facilitate the creation of C++
Python bindings. We present a short study that compares three main
tools, which leads to the method of choice.
%
We present a system that dynamically generates bindings of desired
objects according to user prescriptions, which enables the convenient
use of any subset of bound object types concurrently.  With our system
it is possible to generate a single library that contains bindings for
instances of different \cgal{} class templates, e.g., an instance of
the 2D arrangement class template and an instance of the 2D
triangulation class template, or several libraries when one is
insufficient. If several instances of the same class template must be
bound, several corresponding libraries must be generated, one for each
instance.
%
Finally, the binding themselves are implemented in C++ and their
implementation exploits advanced features in C++. This results with
elegant and compact code that is easy to maintain and extend with
additional bindings.

% =============================================================================
\subsection{Outline}
\label{ssec:intro:outline}
% =============================================================================
The rest of this paper is organized as follows. A study that compares
three main methods for implementing C++ Python bindings is presented
in Section~\ref{sec:comp}. General instructons for dynamically
generating bindings are presented in Section~\ref{sec:bind-gen}.
Description of the bindings generated by our system for various
modules are given in Section~\ref{sec:modules}. Highlighted modern C++
techniques and idioms used in the binding implementation are described
in Section~\ref{sec:bind-code}.

% =============================================================================
\subsection{Conventions}
\label{ssec:intro:conventions}
% =============================================================================
The paper is packaed with code samples written either in the \cmake{},
Python, or C++ languages. We use several conventions to improve
readabilty.  Code excerpt written in C++ is always numberd, while
short code excerpts written in Python is sometimes preceded with
\pyLstinline{>>>}. The C++ identifier \cppLstinline{pb} is an alias
for the namespace \cppLstinline{boost::python}.
